---
title: "ST and VPP workflows"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ST and VPP workflows}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

## Seasonal Trajectories (ST)

### Preview

```{r}
library(TSGenerator)

st_preview <- download_st(
  start = "2020-04-01",
  end = "2020-06-30",
  tile_id = "30TXM",
  product = c("PPI", "QFLAG"),
  download = FALSE
)
```

### Download

```{r}
st_download <- download_st(
  start = "2020-04-01",
  end = "2020-06-30",
  tile_id = "30TXM",
  product = c("PPI", "QFLAG"),
  output_dir = "HRVPP_ST",
  overwrite = FALSE,
  prompt = TRUE
)
```

### Extract PPI

```{r}
st_ts <- extract_ts(
  x = "HRVPP_ST/PPI",
  polygons = parcels,
  id_col = "Parcel_ID",
  fun = "median"
)
```

### Assess temporal completeness

```{r}
plan <- temporal_plan(st_ts, expected_step = 10)
missingness <- summarize_missingness(plan)
completeness <- assess_ts_quality(plan)
```

Do not interpret these completeness classes as ST-QFLAG confidence classes.

### Apply ST-QFLAG when required

```{r}
st_qflag_classes <- quality_info("ST")

st_clean <- mask_quality(
  x = st_raster,
  quality = st_qflag_raster,
  type = "ST",
  min_quality = 4
)
```

The default ST quality policy and any stricter threshold should be reported in the methods of a scientific analysis.

## Vegetation Phenology and Productivity (VPP)

### Preview/download selected parameters

```{r}
vpp_preview <- download_vpp(
  start = "2020-01-01",
  end = "2020-12-31",
  tile_id = "30TXM",
  product = c("SOSD", "MAXD", "EOSD", "LENGTH", "MAXV", "AMPL"),
  season = "s1",
  download = FALSE
)
```

### Product-aware extraction

```{r}
vpp <- extract_vpp(
  x = "HRVPP_VPP",
  polygons = parcels,
  id_col = "Parcel_ID",
  fun = "median"
)
```

The output separates `Date` and numeric `Value`. This avoids treating encoded phenological dates like continuous spectral values.

### VPP quality

```{r}
quality_info("VPP")

vpp_clean <- mask_quality(
  x = vpp_raster,
  quality = vpp_qflag,
  type = "VPP"
)
```

The default VPP mask retains medium/high confidence codes. Use explicit `keep_codes` when the study requires a different policy.

## Missingness modelling

For a table containing `Year`, `DOY`, and the analysed variable:

```{r}
fit <- model_missingness(
  missingness_data,
  year_col = "Year",
  doy_col = "DOY",
  value_col = "Value"
)

plot_missingness(fit, sos = 85, max_doy = 160, eos = 280)
```

The model and plot are separate objects. This makes the statistical result reusable without hidden plotting side effects.
